---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    fig_height: 12
    fig_width: 4
    keep_md: yes
  pdf_document: default
  word_document: default
---


##Loading and preprocessing the data
###1. Load the data (i.e. `read.csv()` )
Since we have forked the instructor's repository into our githubs and cloned it locally, we already have the necessary files.  Simply, change the working directory in R Studio, and then unzip and load the file.

```{r, echo=TRUE}
activitydata<- unzip("activity.zip")
activitydata<- read.csv("../RepData_PeerAssessment1/activity.csv", header=TRUE, sep= ",")
```

###2. Process/transform the data (if necessary) into a format suitable for your analysis
Okay, let's see if we need to fix up the table in anyway using the `str()` function

```{r, echo=FALSE, results="asis"}
str(activitydata)
```

Looks like the date colum has class=FACTOR; we should change that to class=Date. I like working with the **lubridate** package but the 'as.date()' should be fine too.
```{r,echo=TRUE}
install.packages("lubridate")
library(lubridate)
activitydata$date<- ymd(activitydata$date)
```

##What is the mean total number of steps taken per day?
###1. Calculate the total number fo steps taken per day
First, I should calculate the sum for each day, then I can take the mean across the days. I am going to use the **dplyr** package
```{r, echo=TRUE}
groupeddata<- group_by(activitydata, date)
summeddata<- summarise(groupeddata, sum(steps))
names(summeddata)<- c("date", "total steps")
```
or another way to do it is `rowsum(activitydata$steps, activitydata$date)`, where the first argument is the data you want to sum, and the second argument is the 'group_by' argument

###2. Make a histogram of the total number of steps taken each day.
We can use the 'summeddata' dataframe and the base plotting package:
```{r, results = "asis"}
 hist(summeddata$"total steps", breaks=10, main = "Histogram of Total No of Steps by Day")
```

###3. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
summeddata$"total steps"<- as.numeric(as.integer(summeddata$"total steps"))
cleansums<- na.omit(summeddata)
mean<- mean(cleansums$"total steps")
median<- median(cleansums$"total steps")
